Miniproject - Combinatorics

### 2021-12-14 Combinatorics/submission.tex
Added literature search results to combinatorics

    Now consider the sets
    \begin{equation*}
        C'_m := C_j \cup \{ a \} \ \text{where} \ m = \{\underbrace{(1, i, a)}_{\in V_1}, \underbrace{(2, j, b)}_{\in V_2}\} \in E
    \end{equation*}
    These are chains, as $\max C_j < a$ for each $C'_m$.
    Further, for each $x, y \in L_i, i < n$ have that
    \begin{equation*}
        |\{ C'_m \ | \ x \in C'_m \}| = |L_n \times \{ j \ | \ x \in C_j \}| = |L_n \times \{ j \ | \ y \in C_j \}| = |\{ C'_m \ | \ y \in C'_m \}|
    \end{equation*}
    as $M$ is a matching from $V_2$ to $V_1$.
    Finally, for all $x \in L_n$ we have that
    \begin{align*}
        |\{ C'_m \ | \ x \in C'_m \}| =& |\{ m \in M \ | \ \exists i, v \in V_2: m = \{ (i, x), v \} \}| = |\{ (i, x) \ | \ (i, x) \in V_1 \}| = t
    \end{align*}
    as $M$ is a matching from $V_1$ to $V_2$.

@@

\section{Part III}
We even show the slightly stronger stronger statement that the implication to the second condition of Prop.\ref{prop:characterization_lym} already holds ``locally'', i.e. for each layer separately.
\begin{proposition}
    \label{prop:biregular_implies_lym}
    Let $P$ be a graded poset with layers $L_0, ..., L_n$ and let $0 < i \leq n$ such that
    \begin{itemize}
        \item Each element $a \in L_i$ covers the same number of elements in $L_{i - 1}$
        \item Each element $a \in L_{i - 1}$ is covered by the same number of elements in $L_i$
    \end{itemize}
    Then for each $F \subseteq L_i$ have
    \begin{equation*}
        \frac {|\partial F|} {|L_{i - 1}|} \geq \frac {|F|} {|L_i|}
    \end{equation*}
    In particular, if this is true for all $i$, then all the equivalent conditions from Prop.~\ref{prop:characterization_lym} follow.
\end{proposition}
\begin{proof}
    Assume that for all $a \in L_i$
    \begin{equation*}
        |\{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \}| = k \in \N
    \end{equation*}
    and for all $b \in L_{i - 1}$
    \begin{equation*}
        |\{ a \in L_i \ | \ \text{$a$ covers $b$} \}| = l \in \N
    \end{equation*}
    Double-counting pairs $(a, b) \in L_i \times L_{i - 1}$ such that $a$ covers $b$ yields $k|L_i| = l|L_{i - 1}|$.
    Now consider $F \subseteq L_i$.
    Note that by definition of $\partial F$, we have for all $a \in F$ that
    \begin{equation*}
        \{ b \in \partial F \ | \ \text{$a$ covers $b$} \} = \{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \} = \partial \{a\} \subseteq \partial F
    \end{equation*}
    Again by double-counting pairs $(a, b) \in F \times \partial F$ such that $a$ covers $b$, we find
    \begin{align*}
        k|F| =& \sum_{a \in F} k = \sum_{a \in F} |\{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \}| = \sum_{a \in F} |\{ b \in \partial F \ | \ \text{$a$ covers $b$} \}| \\
        =& |\{ (a, b) \in F \times \partial F \ | \ \text{$a$ covers $b$} \}| = \sum_{b \in \partial F} |\{ a \in F \ | \ \text{$a$ covers $b$}\}| \\
        \leq& \sum_{b \in \partial F} |\{ a \in L_i \ | \ \text{$a$ covers $b$}\}| = \sum_{b \in \partial F} l = l |\partial F|
    \end{align*}
    Hence
    \begin{equation*}
        \frac {|\partial F|} {|L_{i - 1}|} = \frac {l |\partial F|} {l |L_{i - 1}|} \leq \frac {k |F|} {k |L_i|} = \frac {|F|} {|L_i|}
    \end{equation*}
\end{proof}

\section{Part IV}
\begin{example}
    Let $\Pi_m = \{ \pi \subseteq \powerset(n) \ | \ \text{$\pi$ partition} \}$ be the poset of partitions ordered by refinement.
    Then $\Pi_m$ is a graded poset with rank function
    \begin{equation*}
        r: \Pi_m \to \{ 0, ..., n - 1 \}, \quad \pi \mapsto n - |\pi|
    \end{equation*}
\end{example}
\begin{proof}
    First of all, have $1 \leq |X| \leq n$ for all subsets $X \subseteq \powerset(n)$, hence the function $r$ is well-defined.
    Next we show that for all $x < y$ have $r(x) < r(y)$.
    However, if $x$ is a proper refinement of $y$, then clearly $|x| > |y|$, so $r(x) < r(y)$.

    Now assume there is a maximal chain $\pi_1 < ... < \pi_t$ in $\Pi_m$.
    As the chain is maximal and $\Pi_m$ has the largest element $[n]$, we may assume that $\pi_t = [n]$ and thus $r(\pi_t) = n - 1$.
    Similarly, we may assume that $p_1 = \powerset(n)$ and so $r(\pi_1) = 0$.
    We have to show that $t = n$, and so it suffices to show that $r(\pi_i) = 1 + r(\pi_{i - 1})$ for all $0 < i \leq t$.

@@

        Using this, we find
        \begin{align*}
            |\partial \mathcal{A}| =& \frac 1 2 |\mathcal{G}| = \frac 1 2 \sum_{A \in [m]^{(m/2)}} |\{ (B, C) \ | \ \{ B, C \} \ \text{partition of} \ [m] \setminus A \}| \\
            =& \frac 1 2 \sum_{A \in [m]^{(m/2)}} 2 S(m/2, 2) = \sum_{A \in [m]^{(m/2)}} 2^{m/2 - 1} - 1 = {m \choose m/2} \left( 2^{m/2 - 1} - 1 \right)
        \end{align*}
    \end{proof}
    We can now plug this into the second condition of Prop.~\ref{prop:characterization_lym} to see
    \begin{align*}
        \frac {|\partial \mathcal{A}|} {|L_{n - 3}|} &= \frac {{m \choose m/2}\left( 2^{m/2 - 1} - 1\right)} {S(m, 3)} = \frac {{m \choose m/2}\left( 2^{m/2 - 1} - 1\right)} {\frac 1 2 \left( 3^{m - 1} - 2^m + 1 \right)} \\
        &\quad \sim \frac {2^{m + m/2}} {\sqrt{n}(3^{m - 1} - 2^m)} \sim \frac {3 \left( 2^{3/2} \right)^m} {\sqrt{n} 3^m} = \frac 3 {\sqrt{n}} c^m
    \end{align*}
    and
    \begin{align*}
        \frac {|\mathcal{A}|} {|L_{n - 2}|} &= \frac {\frac 1 2 {m \choose m/2}} {S(m, 2)} = \frac {{m \choose m/2}} {2^m - 2} \sim \frac {2^m} {\sqrt{n}2^m} = \frac 1 {\sqrt{n}}
    \end{align*}
    for some $0 < c < 1$ (here $\sim$ means asymptotically equivalent as $m \to \infty$).
    In particular, find for sufficiently large $m$ that
    \begin{equation*}
        \frac {|\partial \mathcal{A}|} {|L_{n - 3}|} < \frac {|\mathcal{A}|} {|L_{n - 2}|}
    \end{equation*}
    and so the conditions of Prop.~\ref{prop:characterization_lym} are not satisfied.
\end{example}

@@

    Assume not, then 
    \begin{equation*}
        r(a_i) - r(a_{i - 1}) = \sum_j a_{ij} - a_{(i - 1)j} \geq 2
    \end{equation*}
    For all $j$ we have $a_{ij} \geq a_{(i - 1)j}$ and so there is $s \leq d$ with $a_{is} - a_{(i - 1)s} \geq 2$ or there are different $r, s \leq d$ with $a_{ir} - a_{(i - 1)r}, a_{is} - a_{(i - 1)s} \geq 1$.
    In both cases, find
    \begin{equation*}
        \tilde{a} \in \mathcal{P}_{k, d} \ \text{defined by} \ \tilde{a}_j = \begin{cases}
            a_{(i - 1)j} & \text{if $j \neq s$} \\
            a_{(i - 1)j} + 1 & \text{otherwise}
        \end{cases}
    \end{equation*}
    with $a_{i - 1} < \tilde{a} < a_i$. However, this gives a longer chain
    \begin{equation*}
        a_1 < ... < a_{i - 1} < \tilde{a} < a_i < ... < a_t
    \end{equation*}
    which contradicts the assumed maximality. The claim follows.
\end{proof}
\begin{definition}
    Let $P$ be a graded poset with layers $L_0, ..., L_n$.
    A symmetric chain in $P$ is a chain $C \subseteq P$ such that there is $i \leq n/2$ with
    \begin{equation*}
        \forall 0 \leq j \leq n: \ C \cap L_j \neq \emptyset \ \Leftrightarrow \ i \leq j \leq n - i
    \end{equation*}
    A decomposition $\mathcal{C}$ of $P$ into chains is called symmetric chain decomposition, if all $C \in \mathcal{C}$ are symmetric.
\end{definition}
Note that this definition is compatible with the definition of a symmetric chain for the classical graded poset $\powerset(n)$.
\begin{proposition}
    Let $k, d \geq 1$. Then $\mathcal{P}_{k, d}$ has a symmetric chain decomposition.
\end{proposition}
\begin{proof}
    Use induction on $d$.
    The base case $d = 1$ is trivial, so assume that $d > 1$.
    By induction hypothesis, have a symmetric chain decomposition $\mathcal{C}$ of $\mathcal{P}_{k, d - 1}$.
    For some $C = \{ A_i, A_{i + 1}, ..., A_{d(k - 1) - i} \} \in \mathcal{C}, i < \frac n 2$ with $r(A_j) = j$ and $0 \leq l \leq k, n - 2i - 1$ consider now the chain
    \begin{equation*}
        \tilde{C}_l := \{ (A_{i + l}, j) \ | \ 0 \leq j \leq k - l \} \cup \{ (A_j, k - l) \ | \ i + l < j \leq d(k - 1) - (i + l) \}
    \end{equation*}
    Then $\tilde{C}_l$ is symmetric w.r.t. $i + l$, as for $i + l \leq j \leq dk - (i + l)$ we have
    \begin{align*}
        i + l \leq j \leq k + i \quad &\Rightarrow \quad (A_{i + l}, j - i - l) \in \tilde{C}_l, \\
        k + i < j \leq dk - (i + l) \quad &\Rightarrow \quad (A_{j - k}, k - l) \in \tilde{C}_l
    \end{align*}
    Clearly the $\tilde{C}_l$ are chains and disjoint.
    Hence, we find a symmetric chain decomposition
    \begin{equation*}
        \tilde{\mathcal{C}} := \Bigl\{ \tilde{C}_l \ \Bigm| \ C \in \mathcal{C} \ \text{symmetric chain w.r.t. $i$}, \ 0 \leq l \leq \min\{ k, n - 2i - 1 \} \Bigr\}
    \end{equation*}
    of $\mathcal{P}_{k, d}$.
\end{proof}
Note that for $k = 1$, this is exactly the proof that was done in the lecture to show that $\powerset(d)$ has a symmetric chain decomposition.

@@

    Now consider a (finite) sequence of maximal chains $(C_k)_{k \in I}$ such that the chain corresponding to $u_i < v_j$ occurs $a_{ij} |I|$ times.
    By the assumption, we know that for each $i$
    \begin{equation*}
        |\{k \ | \ u_i \in C_k\}| = \sum_{u_i < v_j} a_{ij}|I| = \frac 1 {|L_0|} |I|
    \end{equation*}
    and so each element $u_i \in L_{l - 1}$ is contained in the same number $|I| / |L_0|$ of chains $C_k$.

@@

My successful-looking attempt is based on the observation that elements of $\mathcal{P}_{k, d}$ ``behave similarly'' (in the sense that there is an element of the automorphism group $\mathrm{Aut}(\mathcal{P}_{k, d})$ mapping one to the other) if the have the same number of equally sized entries.
This looks promising, as the proof for $\powerset(n)$ and equivalently the proof of Prop.~\ref{prop:biregular_implies_lym} relies on the fact that all elements in one layer of $\powerset(n)$ ``behave similarly'' in this sense.

This is formalized in the next definition.
\begin{definition}
    Let $P = \mathcal{P}_{k, d}$. Then define
    \begin{equation*}
        P(n_0, ..., n_k) := \{ x \in P \ | \ |\{i \ | \ x(i) = j\}| = n_j \}
    \end{equation*}
    Further, for any subset $F \subseteq P$ write
    \begin{equation*}
        F(n_0, ..., n_k) := F \cap P(n_0, ..., n_k)
    \end{equation*}
\end{definition}
\begin{lemma}
    \label{prop:properties_n_distribution_subsets}
    The size of $|P(n_0, ..., n_k)|$ is given by the multinomial coefficient
    \begin{equation*}
        |P(n_0, ..., n_k)| = { d \choose n_0, ..., n_k } =  \frac {n!} {n_0! ... n_k!}
    \end{equation*}
    if $n_0 + ... + n_k = d$ (otherwise, it is clearly zero).

    Further, for all $n_0, ..., n_k \in \Z$ with $\sum n_i = d$ and all $a \in P(n_0, ..., n_k), 0 \leq \delta < k$ have
    \begin{align*}
        |\partial a \cap P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k)| &= n_{\delta + 1}, \\
        |\partial^+ a \cap P(n_0, ..., n_\delta - 1, n_{\delta + 1} + 1, ..., n_k)| &= n_\delta
    \end{align*}
    even if any of the $n_i$ are negative.
\end{lemma}
\begin{definition}
    Define the \emph{core} of $\mathcal{P}_{k, d}$ as the poset
    \begin{equation*}
        \mathcal{C}_{k, d} := \{ (n_0, ..., n_k) \ | \ n_0 + ... + n_k = d \}
    \end{equation*}
    whose partial order is the transitive closure of the relation $\prec$ defined by
    \begin{align*}
        (n_0, ..., n_k) \prec (m_0, ..., m_k) \ :\Leftrightarrow \ \exists 0 \leq \delta < k: \ &m_\delta = n_\delta - 1, \ m_{\delta + 1} = n_{\delta + 1} - 1, \\
        &\forall i \neq \delta, \delta + 1: m_i = n_i 
    \end{align*}
\end{definition}
\begin{lemma}
    The layers of $\mathcal{C}_{k, d}$ are
    \begin{equation*}
        L_i = \{ (n_0, ..., n_k) \ | \ \sum_j j n_j = i \}
    \end{equation*}
    Further, we have a surjective map
    \begin{equation*}
        R: \mathcal{P}_{k, d} \to \mathcal{C}_{k, d}, \quad A \mapsto \bigl( |\{j \ | \ A(j) = i\}| \bigr)_{0 \leq i \leq k}
    \end{equation*}
    that preserves the ordering and the fiber of a point $(n_0, ..., n_k)$ is exactly $P(n_0, ..., n_k)$.
\end{lemma}

@@

>>>>

    Now consider the sets
    \begin{equation*}
        C'_m := C_j \cup \{ a \} \ \text{for each} \ m = \{\underbrace{(1, i, a)}_{\in V_1}, \underbrace{(2, j, b)}_{\in V_2}\} \in M
    \end{equation*}
    These are chains, as $\max C_j < a$ for each $C'_m$.
    Further, for each $x, y \in L_i, i < n$ have that
    \begin{equation*}
        |\{ m \ | \ x \in C'_m \}| = |L_n \times \{ j \ | \ x \in C_j \}| = |L_n \times \{ j \ | \ y \in C_j \}| = |\{ C'_m \ | \ y \in C'_m \}|
    \end{equation*}
    as $M$ is a matching from $V_2$ to $V_1$.
    Finally, for all $x \in L_n$ we have that
    \begin{align*}
        |\{ C'_m \ | \ x \in C'_m \}| =& |\{ m \in M \ | \ \exists i, v \in V_2: m = \{ (i, x), v \} \}| = |\{ (i, x) \ | \ (i, x) \in V_1 \}| = t
    \end{align*}
    as $M$ is a matching from $V_1$ to $V_2$.

@@

\section{Part III}
In this part, we show that sufficiently symmetric posets satisfy the conditions introduced in the previous part.
We even show a slightly stronger stronger statement than required in the task, namely that the implication to the second condition of Prop.\ref{prop:characterization_lym} already holds ``locally'', i.e. for each layer separately.
\begin{proposition}
    \label{prop:biregular_implies_lym}
    Let $P$ be a graded poset with layers $L_0, ..., L_n$ and let $0 < i \leq n$ such that
    \begin{itemize}
        \item Each element $a \in L_i$ covers the same number of elements in $L_{i - 1}$
        \item Each element $a \in L_{i - 1}$ is covered by the same number of elements in $L_i$
    \end{itemize}
    Then for each $F \subseteq L_i$ have
    \begin{equation*}
        \frac {|\partial F|} {|L_{i - 1}|} \geq \frac {|F|} {|L_i|}
    \end{equation*}
    In particular, if this is true for all $i$, then all the equivalent conditions from Prop.~\ref{prop:characterization_lym} follow.
\end{proposition}
\begin{proof}
    Assume that for all $a \in L_i$
    \begin{equation*}
        |\{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \}| = k \in \N
    \end{equation*}
    and for all $b \in L_{i - 1}$
    \begin{equation*}
        |\{ a \in L_i \ | \ \text{$a$ covers $b$} \}| = l \in \N
    \end{equation*}
    Double-counting pairs $(a, b) \in L_i \times L_{i - 1}$ such that $a$ covers $b$ yields $k|L_i| = l|L_{i - 1}|$.
    Now consider $F \subseteq L_i$.
    Note that by definition of $\partial F$, we have for all $a \in F$ that
    \begin{equation*}
        \{ b \in \partial F \ | \ \text{$a$ covers $b$} \} = \{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \} = \partial \{a\} \subseteq \partial F
    \end{equation*}
    Again by double-counting pairs $(a, b) \in F \times \partial F$ such that $a$ covers $b$, we find
    \begin{align*}
        k|F| =& \sum_{a \in F} k = \sum_{a \in F} |\{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \}| = \sum_{a \in F} |\{ b \in \partial F \ | \ \text{$a$ covers $b$} \}| \\
        =& |\{ (a, b) \in F \times \partial F \ | \ \text{$a$ covers $b$} \}| = \sum_{b \in \partial F} |\{ a \in F \ | \ \text{$a$ covers $b$}\}| \\
        \leq& \sum_{b \in \partial F} |\{ a \in L_i \ | \ \text{$a$ covers $b$}\}| = \sum_{b \in \partial F} l = l |\partial F|
    \end{align*}
    Hence
    \begin{equation*}
        \frac {|\partial F|} {|L_{i - 1}|} = \frac {l |\partial F|} {l |L_{i - 1}|} \leq \frac {k |F|} {k |L_i|} = \frac {|F|} {|L_i|}
    \end{equation*}
\end{proof}

\section{Part IV}
\begin{example}
    Let $\Pi_m = \{ \pi \subseteq \powerset(m) \ | \ \text{$\pi$ partition} \}$ be the poset of partitions ordered by refinement.
    Then $\Pi_m$ is a graded poset with rank function
    \begin{equation*}
        r: \Pi_m \to \{ 0, ..., m - 1 \}, \quad \pi \mapsto m - |\pi|
    \end{equation*}
\end{example}
\begin{proof}
    First of all, have $1 \leq |X| \leq n$ for all subsets $X \subseteq \powerset(n)$, hence the function $r$ is well-defined.
    Next we show that for all $x < y$ have $r(x) < r(y)$.
    However, if $x$ is a proper refinement of $y$, then clearly $|x| > |y|$, so $r(x) < r(y)$.

    Now assume there is a maximal chain $\pi_1 < ... < \pi_t$ in $\Pi_m$, and want to show that $t = m$.
    As the chain is maximal and $\Pi_m$ has the largest element $[m]$, we may assume that $\pi_t = [m]$ and thus $r(\pi_t) = m - 1$.
    Similarly, we may assume that $p_1 = \{ \{a\} \ | \ a \in [m]\}$ and so $r(\pi_1) = 0$.
    So it suffices to show that $r(\pi_i) = 1 + r(\pi_{i - 1})$ for all $0 < i \leq t$.

@@

        Using this, we find
        \begin{align*}
            |\partial \mathcal{A}| =& \frac 1 2 |\mathcal{G}| = \frac 1 2 \sum_{A \in [m]^{(m/2)}} |\{ (B, C) \ | \ \{ B, C \} \ \text{partition of} \ [m] \setminus A \}| \\
            =& \frac 1 2 \sum_{A \in [m]^{(m/2)}} 2 S(m/2, 2) = \sum_{A \in [m]^{(m/2)}} 2^{m/2 - 1} - 1 = {m \choose m/2} \left( 2^{m/2 - 1} - 1 \right)
        \end{align*}
    \end{proof}
    We can now plug this into the second condition of Prop.~\ref{prop:characterization_lym} to see
    \begin{align*}
        \frac {|\partial \mathcal{A}|} {|L_{m - 3}|} &= \frac {{m \choose m/2}\left( 2^{m/2 - 1} - 1\right)} {S(m, 3)} = \frac {{m \choose m/2}\left( 2^{m/2 - 1} - 1\right)} {\frac 1 2 \left( 3^{m - 1} - 2^m + 1 \right)} \\
        &\quad \sim \frac {2^{m + m/2}} {\sqrt{m}(3^{m - 1} - 2^m)} \sim \frac {3 \left( 2^{3/2} \right)^m} {\sqrt{m} 3^m} = \frac 3 {\sqrt{m}} c^m
    \end{align*}
    and
    \begin{align*}
        \frac {|\mathcal{A}|} {|L_{m - 2}|} &= \frac {\frac 1 2 {m \choose m/2}} {S(m, 2)} = \frac {{m \choose m/2}} {2^m - 2} \sim \frac {2^m} {\sqrt{m}2^m} = \frac 1 {\sqrt{m}}
    \end{align*}
    for some $0 < c < 1$ (here $\sim$ means asymptotically equivalent as $m \to \infty$).
    In particular, find for all $\epsilon > 0$ and sufficiently large $m$ that
    \begin{equation*}
        \frac {|\partial \mathcal{A}|} {|L_{m - 3}|} < \epsilon \frac {|\mathcal{A}|} {|L_{m - 2}|}
    \end{equation*}
    and so the conditions of Prop.~\ref{prop:characterization_lym} are not satisfied.
\end{example}

@@

    Assume not, then 
    \begin{equation*}
        r(a_i) - r(a_{i - 1}) = \sum_j a_{ij} - a_{(i - 1)j} \geq 2
    \end{equation*}
    For all $j$ we have $a_{ij} \geq a_{(i - 1)j}$ and so there is $s \leq d$ with $a_{is} - a_{(i - 1)s} \geq 2$ or there are different $r, s \leq d$ with $a_{ir} - a_{(i - 1)r}, a_{is} - a_{(i - 1)s} \geq 1$.
    In both cases, find
    \begin{equation*}
        \tilde{a} \in \mathcal{P}_{k, d} \ \text{defined by} \ \tilde{a}_j = \begin{cases}
            a_{(i - 1)j} & \text{if $j \neq s$} \\
            a_{(i - 1)j} + 1 & \text{otherwise}
        \end{cases}
    \end{equation*}
    with $a_{i - 1} < \tilde{a} < a_i$. However, this gives a longer chain
    \begin{equation*}
        a_1 < ... < a_{i - 1} < \tilde{a} < a_i < ... < a_t
    \end{equation*}
    which contradicts the assumed maximality. The claim follows.
\end{proof}
\begin{definition}
    Let $P$ be a graded poset with layers $L_0, ..., L_n$.
    A symmetric chain in $P$ is a chain $C \subseteq P$ such that there is $i \leq n/2$ with
    \begin{equation*}
        \forall 0 \leq j \leq n: \ C \cap L_j \neq \emptyset \ \Leftrightarrow \ i \leq j \leq n - i
    \end{equation*}
    A decomposition $\mathcal{C}$ of $P$ into chains is called symmetric chain decomposition, if all $C \in \mathcal{C}$ are symmetric.
\end{definition}
Note that this definition is compatible with the definition of a symmetric chain for the classical graded poset $\powerset(n)$.
\begin{proposition}
    Let $k, d \geq 1$. Then $\mathcal{P}_{k, d}$ has a symmetric chain decomposition.
\end{proposition}
\begin{proof}
    Use induction on $d$.
    The base case $d = 1$ is trivial, so assume that $d > 1$.
    By induction hypothesis, have a symmetric chain decomposition $\mathcal{C}$ of $\mathcal{P}_{k, d - 1}$.
    Now consider some $C = \{ A_i, A_{i + 1}, ..., A_{d(k - 1) - i} \} \in \mathcal{C}, i < \frac n 2$ with $r(A_j) = j$.
    For each $0 \leq l \leq k, n - 2i - 1$ define the chain
    \begin{equation*}
        \tilde{C}_l := \{ \underbrace{(A_{i + l}, j)}_{\mathclap{:= (a_1, ..., a_{d - 1}, j) \ \text{if} \ A_{i + l} = (a_1, ..., a_{d - 1})}} \ | \ 0 \leq j \leq k - l \} \cup \{ (A_j, k - l) \ | \ i + l < j \leq d(k - 1) - (i + l) \}
    \end{equation*}
    Then $\tilde{C}_l$ is symmetric w.r.t. $i + l$, as for $i + l \leq j \leq dk - (i + l)$ we have
    \begin{align*}
        i + l \leq j \leq k + i \quad &\Rightarrow \quad (A_{i + l}, j - i - l) \in \tilde{C}_l, \\
        k + i < j \leq dk - (i + l) \quad &\Rightarrow \quad (A_{j - k}, k - l) \in \tilde{C}_l
    \end{align*}
    Clearly the $\tilde{C}_l$ are chains and disjoint.
    Hence, we find a symmetric chain decomposition
    \begin{equation*}
        \tilde{\mathcal{C}} := \Bigl\{ \tilde{C}_l \ \Bigm| \ C \in \mathcal{C} \ \text{symmetric chain w.r.t. $i$}, \ 0 \leq l \leq \min\{ k, n - 2i - 1 \} \Bigr\}
    \end{equation*}
    of $\mathcal{P}_{k, d}$.
\end{proof}
Note that for $k = 1$, this is exactly the proof that was done in the lecture to show that $\powerset(d)$ has a symmetric chain decomposition.

@@

    Now consider a (finite) sequence of maximal chains $(C_k)_{k \in I}$ such that the chain corresponding to $u_i < v_j$ occurs $a_{ij} |I|$ times.
    By the assumption, we know that for each $i$ have
    \begin{equation*}
        |\{k \ | \ u_i \in C_k\}| = \sum_{u_i < v_j} a_{ij}|I| = \frac 1 {|L_0|} |I|
    \end{equation*}
    and so each element $u_i \in L_{l - 1}$ is contained in the same number $|I| / |L_0|$ of chains $C_k$.

@@

My successful-looking attempt at proving this is based on the observation that elements of $\mathcal{P}_{k, d}$ ``behave similarly'' (in the sense that there is an element of the automorphism group $\mathrm{Aut}(\mathcal{P}_{k, d})$ mapping one to the other) if the have the same number of equally sized entries.
This looks promising, as the proof for $\powerset(n)$ relies on the fact that all elements in one layer of $\powerset(n)$ ``behave similarly'' in this sense.
In particular, note that Prop.~\ref{prop:biregular_implies_lym} would already be applicable in such a case.

This is formalized in the next definition.
\begin{definition}
    Let $P = \mathcal{P}_{k, d}$. Then define
    \begin{equation*}
        P(n_0, ..., n_k) := \bigl\{ x \in P \ \bigm| \ |\{i \ | \ x(i) = j\}| = n_j \bigl\}
    \end{equation*}
    Further, for any subset $F \subseteq P$ write
    \begin{equation*}
        F(n_0, ..., n_k) := F \cap P(n_0, ..., n_k)
    \end{equation*}
\end{definition}
\begin{lemma}
    \label{prop:properties_n_distribution_subsets}
    The size of $|P(n_0, ..., n_k)|$ is given by the multinomial coefficient
    \begin{equation*}
        |P(n_0, ..., n_k)| = { d \choose n_0, ..., n_k } =  \frac {d!} {n_0! ... n_k!}
    \end{equation*}
    if $n_0 + ... + n_k = d$ and $n_i \geq 0$ (otherwise, it is clearly zero).

    Further, for all $n_0, ..., n_k \in \Z$ with $\sum n_i = d$ and all $a \in P(n_0, ..., n_k), 0 \leq \delta < k$ have
    \begin{align*}
        |\partial a \cap P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k)| &= n_{\delta + 1}, \\
        |\partial^+ a \cap P(n_0, ..., n_\delta - 1, n_{\delta + 1} + 1, ..., n_k)| &= n_\delta
    \end{align*}
    even if any of the $n_i$ are negative.
\end{lemma}
\begin{proof}
    For $n_0 + ... + n_k = d$ and $n_i \geq 0$ we have
    \begin{align*}
        &|P(n_0, ..., n_k)| = |\bigl\{ x \in P \ \bigm| \ |\{i \ | \ x(i) = j\}| = n_j\bigl\}| \\
        =& \sum_{A_0 \in [d]^{(n_0)}} |\bigl\{ x \in \{0, ..., k\}^{[d] \setminus A_0} \ \bigm| \ |\{i \ | \ x(i) = j\}| = n_j\bigl\}| \\
        =& \sum_{A_0 \in [d]^{(n_0)}} \sum_{A_1 \in ([d] \setminus A_0)^{(n_1)}} |\bigl\{ x \in \{0, ..., k\}^{[d] \setminus A_0 \setminus A_1} \ \bigm| \ |\{i \ | \ x(i) = j\}| = n_j\bigl\}| \\
        =& \ ... \\
        =& \sum_{A_0 \in [d]^{(n_0)}} ... \sum_{A_k \in ([d] \setminus A_0 \setminus ... \setminus A_{k - 1})^{(n_k)}} 1 \\
        =& \sum_{A_0 \in [d]^{(n_0)}} ... \sum_{A_{k - 1} \in ([d] \setminus A_0 \setminus ... \setminus A_{k - 2})^{(n_k)}} {d - n_0 - ... - n_{k - 1} \choose n_k} \\
        =& \ ... \\
        =& {d \choose n_0} {d - n_0 \choose n_1} ... {d - n_0 - ... - n_{k - 1} \choose n_k} = \prod_i \frac {(d - n_0 - ... - n_{i - 1})!} {n_i! (d - n_0 - ... - n_i!)} \\
        =& \frac {d!} {n_0! ... n_k!} \prod_i \frac {(d - n_0 - ... - n_i)!} {(d - n_0 - ... - n_i)!} = \frac {d!} {n_0! ... n_k!}
    \end{align*}
    Further have for all $n_0 + ... + n_k = d$ and $a \in P(n_0, ..., n_k)$ that there is a bijection
    \begin{align*}
        \{ i \in [d] \ | \ a(i) = \delta + 1 \} &\to \partial a \cap P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k), \\
        i &\mapsto f_{a, i}: j \mapsto \begin{cases}
            a(j) & \text{if $j \neq i$} \\
            a(j) - 1 & \text{otherwise}
        \end{cases}
    \end{align*}
    It is easy to see that the image of some $i$ is indeed an element of $\partial a$ and of $P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k)$.
    Clearly the function is also injective, and surjectivity can be shown by considering some $b \in \partial a \cap P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k)$ and showing by induction that
    \begin{equation*}
        \{ i \ | \ b(i) = j \} = \{ i \ | \ a(i) = j \}
    \end{equation*}
    for all $j \neq \delta, \delta + 1$.
    Note that the set $\{ i \in [d] \ | \ a(i) = \delta + 1 \}$ has size exactly $n_{\delta + 1}$ and so the claim follows.

    The analogue statement for the upper shadow of $a$ can be proven in the same way.
\end{proof}
\begin{definition}
    Define the \emph{core} of $\mathcal{P}_{k, d}$ as the poset
    \begin{equation*}
        \mathcal{C}_{k, d} := \{ (n_0, ..., n_k) \ | \ n_0 + ... + n_k = d \}
    \end{equation*}
    whose partial order is the transitive closure of the relation $\prec$ defined by
    \begin{align*}
        (n_0, ..., n_k) \prec (m_0, ..., m_k) \ :\Leftrightarrow \ \exists 0 \leq \delta < k: \ &m_\delta = n_\delta - 1, \ m_{\delta + 1} = n_{\delta + 1} - 1, \\
        &\forall i \neq \delta, \delta + 1: m_i = n_i 
    \end{align*}
\end{definition}
\begin{lemma}
    The layers of $\mathcal{C}_{k, d}$ are
    \begin{equation*}
        L_i = \{ (n_0, ..., n_k) \ | \ \sum_j j n_j = i \}
    \end{equation*}
    Further, we have a surjective map
    \begin{equation*}
        R: \mathcal{P}_{k, d} \to \mathcal{C}_{k, d}, \quad A \mapsto \bigl( |\{j \ | \ A(j) = i\}| \bigr)_{0 \leq i \leq k}
    \end{equation*}
    that preserves the ordering and the fibre of a point $(n_0, ..., n_k)$ is exactly $P(n_0, ..., n_k)$.
\end{lemma}
\begin{proof}
    All of these statements are very easy to check.
\end{proof}

@@

After quite a long time trying to prove that $\mathcal{P}_{k, d}$ satisfies an LYM-inequality analogue, I finally gave up and searched the literature.
There I found that $\mathcal{P}_{k, d}$ is also called divisor lattice and indeed we have the following statement.
\begin{proposition}
    $\mathcal{P}_{k, d}$ satisfies the conditions of Prop.~\ref{prop:characterization_lym}, i.e. an LYM-inequality.
\end{proposition}
\begin{proof}
    See e.g. \cite[Theorem 4.2.3]{anderson}.
\end{proof}
Interestingly, I have also found a variant of the LYM inequality for the poset $\Pi_m$ that we studied earlier.
\begin{proposition}
    Let $\mathcal{A} \subseteq \Pi_m$ be a nonempty antichain. Then define
    \begin{equation*}
        a_{n_1, ..., n_k} := |\bigl\{ A \in \mathcal{A} \ \bigm| \ |A| = k, \ |\{ S \in A \ | \ |S| = j \}| = n_j \bigr\}|
    \end{equation*}
    the number of partitions $A \in \mathcal{A}$ such that the classes in $A$ have sizes $n_1, ..., n_k$.
    Then
    \begin{equation*}
        \sum_{k \geq 1} \sum_{n_1, ..., n_k} \frac {a_{n_1, ..., n_k} k!} {{m \choose n_1, ..., n_k} {m - 1 \choose k - 1}} \leq 1
    \end{equation*}
\end{proposition}
\begin{proof}
    See \cite{lym_partitions}.
\end{proof}
